### 2023-02-21 Conference Submission/DraftDoc.tex
more edits from giulio

\begin{section}{Introduction}
The modelling of temporal networks is an important task in many real world applications XXXRef.
Temporal networks can be seen as dynamical systems: XXX.
Discovering the underlying equations governing these dynamical systems proves is challenging, because XXX.
Here, we proposes an hybrid statistical and deep learning framework that allows to model temporal networks as continous-time dynamical system, discover a fitting set of differential equations describing it, and, exploiting that discovery, predict the time evolution of a network.

XXX Nodes within the network influence each other and XXX so we might wish to draw on the rich mathematical literature of differential equation modelling.

@@

Here, we overcome the discretness problem by interpreting networks as Random Dot Product Graphs; a well established statistical model for complex networks, that embeds nodes in a low-dimensional metric space[citation]. In this way we translate the hard problem of modelling discrete events in the space of networks to the easier problem of modelling continous change in the embedding space. Because the progression of the system can now be viewed as a continuous transformation, we can now look to model the change of this system. To regenerate the network from the embedding, we can simply multiply our embedding matrices together. This will give us a random dot product graph, where the entries (i,j) of the matrix will be the probability of an edge from i to j existing[citation]. In this form, the problem can be viewed as a dynamical system We believe this approach will be useful when applied to areas of medicine, especially protein interaction networks; population dynamics for network ecology; and social network modelling.

By using this process, we also open the possibility of using the endlessly flexible toolbox of machine learning. In this paper we explore the usefulness of neural ODEs, a neural network that is trained to differential equation of temporal data, for discovering the governing function of a set of simple, synthetic temporal networks. This results in a black box DE, which is not very helpful for understanding our systems. To give more insight into what is really going on, we then use symbolic regression to find a function whose output is similar to that of the neural network. By doing this we may further understand the important factors that govern our temporal network.

@@

\begin{section}{Methodology}
    We used a singular value decomposition of each matrix to embed our network in a low dimensional space.\
        \[
        A = U\Sigma V'
        \]
    Where $\Sigma_{ii} = \sigma_i$ and $\sigma_i > sigma_{i+1}$ and each $\sigma_i$ is a non-negative square root of the matrices $AA', \space A'A$. $U$ is a matrix made from the orthonormalised eigenvectors of $AA'$ associated with $\sigma_i$, and $V$ is a matrix made from the orthonormalised eigenvectors of $A'A$ associated with $\sigma_i$. Each row $i$ of $U$ and $V$ represents the in and out position of node $i$ respectively. This can be truncated at an arbitrary number of eigenpairs to form a low dimentional approximation to $A$.
        \[
        A \approx \hat U \hat \Sigma \hat V'
        \]\\
    Where $\hat U, \hat \Sigma, \hat V$ are the first $d$ columns of $U, \Sigma, V$ \cite{golub1971singular}. $\hat U \sqrt{\hat \Sigma}$ is still associated with the in edges of our nodes, and $\sqrt{\hat \Sigma} \hat V' $ is still associated with the out edges of our nodes, but now we have simplified our system whilst keeping as much information as possible due to our descending order of singular values. These two matrices form our $d$ dimensional embedding.\\

>>>>

\begin{section}{Introduction}
The modelling of temporal networks is an important task in many real world applications including symptom interactions for mental health, epidemiology, and protein interactions \cite{jordan2020current,contreras2020temporal,lucas2021inferring,jin2009identifying,masuda2013predicting}.
Temporal networks can be seen as dynamical systems: that is a system in which we have points, in our case nodes in a network, whose states, the edges connecting them, that vary dependent in time.
Discovering the underlying equations governing these dynamical systems proves challenging, because changes in network structure are typically observed in the form of discrete jumps from one state to another, for example an edge between two nodes not being observed at the first timestep then appearing at the next.
Here, we proposes a hybrid statistical and deep learning framework that allows us to model temporal networks as continous-time dynamical systems, discover a fitting set of differential equations describing it, and, exploiting that discovery, predict the time evolution of a network.

Differential equations are useful for modelling systems where the state of one variable can effect the trajectories of other variables. We observe this behaviour in temporal networks; nodes' connections within the network can influence the formation and decay of edges between other nodes, for example the phenonmenon of preferential attatchment observed in \cite{newman2001clustering,capocci2006preferential}. With this in mind we might wish to draw on the rich mathematical literature of differential equation modelling.

@@

Here, we overcome the discretness problem by interpreting networks as Random Dot Product Graphs; a well established statistical model for complex networks, that embeds nodes in a low-dimensional metric space\cite{athreya2017statistical}. In this way we translate the hard problem of modelling discrete events in the space of networks to the easier problem of modelling continous change in the embedding space. Then, we define and use systems of Neural Network Differential Equations (NNDE) to approximate the time evolution of the embedding space, and symbolic regression techniques to discover the functional form of the fitted NNDEs. These functional forms are interpretable (as they read as classic differential equations) and allow to predict forward in time the evolution of the temporal networks.

In this manuscript, we prove that the temporal network prediction problem can be successfully re-interpreted as a dynamical system modelling problem. In particular, we apply our proposed framework to three small example temporal networks with the hope of exploring the limitations and strengths of the proposed framework.
The framework we are introducing is extremely flexible, and our research regarding the optimal structure of the Neural Networks used for the NNDEs is just started.
We are confident that future research can identify more fitting Neural Network structures than the simple one adopted here.
For this reason, we did not yet attempt to benchmark our model against other classic temporal network prediction methods.
As it is completely general, we believe that the framework we are introducing can be usefully applied to areas of medicine, especially protein interaction networks; population dynamics for network ecology; and social network modelling. In particular, we discuss how specific domain knowledge relative to the prediction scenario can be taken into account, moving from NNDEs to Universal Differential Equations.

@@

\begin{section}{Methodology}

    XXX A couple of rows about RDPG XXX

    XXX what are these matrix XXX

    We used a Singular Value Decomposition of each matrix to embed our network in a low dimensional space.\
        \[
        A = U\Sigma V'
        \]
    XXX Let's not explain SVD, at most give a ref XXX
    This can be truncated at an arbitrary number of eigenpairs to form a low dimentional approximation to $A$.
        \[
        A \approx \hat U \hat \Sigma \hat V'
        \]\\
    Where $\hat U, \hat \Sigma, \hat V$ are the first $d$ columns of $U, \Sigma, V$ \cite{golub1971singular}. $\hat U \sqrt{\hat \Sigma}$ is still associated with the in edges of our nodes, and $\sqrt{\hat \Sigma} \hat V' $ is still associated with the out edges of our nodes, but now we have simplified our system whilst keeping as much information as possible due to our descending order of singular values. These two matrices form our $d$ dimensional embedding.\\
